\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-27 paper/experiments.tex
paper: pass over experiments section

@@

\subsection{Data sets} \label{subsec:datasets}
\subsubsection{Center for Disease Control: Surveys}
We use survey data collected by the 
Center for Disease Control and Prevention (CDC) in the United States. We
experiment on a set of tables collected as part of the 2013-2014 National
Health and Nutrition Examination Survey (NHANES), a series of studies
conducted by the CDC on a national sample of several thousand individuals\cite{cdc-data}.
The data consists on survey responses, physical examinations, and laboratory
results, amongst others. Our dataset is available for direct download on the CDC's organization site on 
kaggle.com, a popular data science website.

There are a total of 6 tables associated with the NHANES data set. We use three tables:

\begin{itemize}
	\item Demographics: outlines demographic information of subjects
	\item Examinations: physical exam results
	\item Laboratory: laboratory exam results
\end{itemize}

The original tables have a large number of attributes, in some cases providing more granular
tests results or alternative metrics. In order to simplify the presentation and
exploration of queries, we focused on a subset of the attributes for each table.~\Cref{table:nhanes-description}
shows the attributes selected, along with the
percentage of nulls. It is worth noting that for readability, we have replaced the
NHANES variable names with self-explanatory attribute names.

\begin{table}
    \centering
    \begin{subtable}{0.5\textwidth}
        \centering
       \input{tables/cdc_demo}
        \caption{Demographics}
    \end{subtable}
    \begin{subtable}{0.5\textwidth}
        \centering
        \input{tables/cdc_labs}
        \caption{Laboratory Results}
    \end{subtable}
       \begin{subtable}{0.5\textwidth}
        \centering
        \input{tables/cdc_exams}
        \caption{Physical Results}
    \end{subtable}
    \caption{Missing value distribution within each table/attribute in CDC NHANES 2013-2014 data}
    \label{table:nhanes-description} 
\end{table}

All tables used consists of 10 attributes. The laboratory and examinations data,
referred to as \textit{labs} and \textit{exams} in our queries, consist of 9813 rows,
while the demographics data (\textit{demo}) have 10175 rows.

\subsubsection{freeCodeCamp: 2016 New Coder Survey}
For our second set of experiments, we chose to use data collected
by freeCodeCamp as a part of a survey of new software developers
(both professional and amateur)\cite{fcc-data}. freeCodeCamp is an open-source
community that intends to help users learn how to program. In addition,
to providing training material for users, it also pairs users with
nonprofits to develop software solutions. Their \textit{2016 New Coder Survey} consists of responses by over 15,000 people to 48 different
questions. The survey targeted users who were in one way or another
related to coding organizations.

For our purposes, we use a version of the data that has been 
pre-processed, but where many missing values remain. It is
worth noting that given the table design (where many attributes
are related), many of these missing values are to be
expected. The dataset we started with had 15,620 rows and 113
columns. We chose 17 attributes, which are outlined in~\Cref{table:fcc-description}, along with the percentage of missing values.

@@

For our final experiment, we introduce a simple aggregate query over data from the American Community Survey (ACS), which
provides a number of public data sets collected by the U.S. Census Bureau.
We used a cleaned version of the 2012 Public Use Microdata Sample (PUMS) data kindly provided by the authors of~\cite{akande2015empirical}.
Given that the data had been cleaned, we uniformly introduced missing values across 40 percent of the values. The final dataset consists
of 671,153 rows and 37 integer columns.

\subsection{Queries}
We collect a set of queries (\Cref{fig:queries}) that we think are both representative and interesting
to plan. They could reasonably be written by a user in the course of data analysis.

The queries on the CDC NHANES data consist not only of projections and selections,
but also interesting joins and aggregates. Our aim was to craft meaningful queries
that would provide performance figures relevant to practitioners who may use
similar datasets.

>>>>

\newcommand{\demorows}{10175}
\newcommand{\labexrows}{9813}

@@

\subsection{Data sets} \label{subsec:datasets}
\subsubsection{CDC NHANES}
We use survey data collected by the 
Centers for Disease Control and Prevention (CDC) in the United States. We
experiment on a set of tables collected as part of the 2013--2014 National
Health and Nutrition Examination Survey (NHANES), a series of studies
conducted by the CDC on a national sample of several thousand individuals~\cite{cdc-data}.
The data consists on survey responses, physical examinations, and laboratory
results, amongst others.

There are 6 tables in the NHANES data set. We use three tables for our experiments:

\begin{itemize}
	\item \emph{Demographics}: demographic information of subjects
	\item \emph{Examinations}: physical exam results
	\item \emph{Laboratory}: laboratory exam results
\end{itemize}

The original tables have a large number of attributes, in some cases providing more granular tests results or alternative metrics.
We focused on a subset of the attributes for each table to simplify the presentation and exploration of queries.
\Cref{table:nhanes-description} shows the attributes selected, along with the percentage of null values for each attribute.
For readability, we have replaced the NHANES variable names with self-explanatory attribute names.

\begin{table}
  \centering
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_demo}
    \caption{Demographics. \demorows{} rows.}
  \end{subtable}
  \par\medskip
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_labs}
    \caption{Laboratory Results. \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_exams}
    \caption{Physical Results. \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \caption{Missing value distribution for each table/attribute in CDC NHANES 2013--2014 data.}\label{table:nhanes-description} 
\end{table}

\subsubsection{freeCodeCamp 2016 New Coder Survey}
For our second set of experiments, we chose to use data collected
by freeCodeCamp as a part of a survey of new software developers
(both professional and amateur)~\cite{fcc-data}. freeCodeCamp is an open-source
community that helps users learn how to program. Their \textit{2016 New Coder Survey} consists of responses by over 15,000 people to 48 different
demographic and programming-related questions.
The survey targeted users who were related to coding organizations.

We use a version of the data that has been pre-processed, but where missing values remain.
\todobox{Give example of expected missing value}{However, many of the missing values are expected because the data has been de-normalized.}
The original dataset has 15,620 rows and 113 attributes.
We chose a subset of 17 attributes, shown in~\Cref{table:fcc-description}, along with the percentage of missing values.

@@

\subsubsection{American Community Survey}
For our final experiment, we introduce a simple aggregate query over data from the American Community Survey (ACS), which
provides a number of public data sets collected by the U.S. Census Bureau.
We used a cleaned version of the 2012 Public Use Microdata Sample (PUMS) data kindly provided by the authors of~\cite{akande2015empirical}.
Given that the data had been cleaned, we artificially dirtied it by removing 40\% of the values uniformly at random.
The final dataset consists of 671,153 rows and 37 integer columns.

\subsection{Queries}
We collected a set of queries (\Cref{fig:queries}) that we think are both representative and interesting
to plan. We believe that they could reasonably be written by a user in the course of data analysis.

The queries on the CDC NHANES data consist not only of projections and selections, but also interesting joins and aggregates.
Our aim was to craft meaningful queries
that would provide performance figures relevant to practitioners using
similar datasets.
